Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function

### 2022-07-27 Manuscript/manuscript.Rmd
update manuscript with edits for brendan

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
any ecosystem function. This challenge remains because it is difficult to
manipulate microbiomes composition directly, we cannot know a priori which members of the
microbiome limit the rate of an ecosystem function, and microbiomes covary with
the environment. We propose that artificial ecosystem selection
can overcome these challenges. Artificial ecosystem selection involves creating
replicate ecosystems, selecting replicates with a desired ecosystem-level
trait, and using the microbiomes from the selected ecosystems to inoculate a
new set of ecosystems. We performed artificial ecosystem selection on replicate
soil microcosms to select for increased soil methane oxidation rate. We
observed a strong response to selection with a 51% increase in methane oxidation rate per round over five
rounds of selection. This suggests that a large portion of the variation in
methane oxidation rate in these soils can be attributed to microbiome
variation. We also found that selection did not enrich for known methane
oxidizers; rather microbes that responded to selection included `r length(da_fams)`
families such as *Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result
is in contrast to the typical assumption that the rate of an ecosystem function
is limited by the final step in the associated microbial pathway. Our study
demonstrates that variation in microbiome composition can contribute to
variation in the rate of ecosystem function.

@@

We applied artificial ecosystem selection on soil microbiomes by selecting on
the ecosystem-scale CH~4~ oxidation rate.  Using this approach, we were able to
estimate how much of the variation in CH~4~ oxidation rate was associated with
variation in the microbiome.  To identify which microorganisms could be markers
of CH~4~ oxidation rate, we compared the composition of the artificially
selected microbiomes to a control set of microbiomes without selection.  We
then evaluated whether these markers meet our underlying assumptions about
which taxa limit the rate of ecosystem function.

@@

We estimated heritability as the regression of mid-offspring on mid-parent (`r figures("herit", display = "cite")`).
Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
Neutral treatment `r print_lm(herit_model, 2)`. Notably, the sign of the effect
was reversed with a positive heritability for the Positive treatment and a negative heritability for the Neutral treatment.

@@

To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
first looked at which taxa were unique to the Positive and Neutral treatments
aggregated at the family level. Several taxa were present in the Positive
treatment and absent in the Neutral treatment. Most of these families had low
prevalence (present in 3 or fewer samples) and low abundance (median < 5 reads).
There were two families unique to the Positive treatment with relatively high
prevalence. This included an ASV that was a member of the Bacteroidia Class with
no lower taxonomic assignment. This ASV had a prevalence of 10/12 and a median
abundance of ~2 reads. The other prevalent family was a member of the
*Silvanigrellaceae*, a newly described family placed in its own order. This family
was present in all 12 samples and had a median abundance of 8 reads.
*Silvanigrella* is the only cultivated member of
*Silvanigrellaceae* and was isolated from a temperate fresh water lake
[@hahn2017]. Of the families unique to the Neutral treatment, only one had a
prevalence greater than 2/12. This family, *Armatimonadaceae*, was present in half
of the Neutral samples (prevalence = 6/12). The type strain for *Armatimonadaceae* was isolated from
the rhizosphere of *Phragmites australis* [@tamaki2011].

The remaining families were shared between the Positive and Neutral treatment,
but some differed in their relative abundance. To identify taxa that responded
to selection on soil CH~4~ oxidation rate, we tested the differential relative
abundance of families in the Positive jars relative to the Neutral jars within
Passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
families that were enriched or depleted in the Positive treatment relative to
the Neutral treatment (`r figures("da", display = "cite")`).

@@

We did not identify any methanotrophic families in the overall differential abundance analysis. 
However, we wanted to look more closely at the known methanotrophs in our dataset to be sure they did not have an effect.
To do this, we subset all of the ASVs in our dataset that were in families that contained methanotrophs.
Only two families were represented: *Methylacidiphilaceae* and *Beijerinckiaceae*. 
Aggregating reads at the family level, neither family was differentially abundant between the two treatments.
However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the Positive treatment and the genus
*Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the Positive treatment.
While many *Beijerinckiaceae* are methanotrophs, several taxa in this family have lost the the ability to oxidize CH~4~ and it appears that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow on methanol [@tamas2014, @dedysh2016]. Therefore, based on this 16S analysis, it appears that no methanotrophs were enriched in the Positive treatment.

@@

We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome generates variation in the rate of ecosystem function independent of the underlying environmental conditions. While it is well understood that the microbiome mediates a variety of ecosystem functions, a fundamental question in microbial ecology is whether we need to understand variation in the composition of microbiomes to understand variation in the rate of ecosystem functions. To address that question, we performed artificial ecosystem selection on soil microbiomes by selecting for ecosystems with high methane oxidation rates. We found that there was a significant increase in methane oxidation rate as a response to selection. Each generation of soil ecosystems was created from the same pool of sterile soil substrate and a small inoculum from the previous generation. Therefore, it is unlikely that this response to selection was due to changes in the abiotic conditions of the ecosystem. In addition, the response was only observed in the Positive treatment and not in the Neutral treatment, which shows that this was not a side effect of passaging in jars. Therefore, we conclude that variation in the microbiome can generate variation in ecosystem functions independent of the environment.

@@

Notably, the Neutral selection treatment displayed a negative heritability of `r myround(coefficients(herit_model)[2], 2)`. Negative heritability estimates can arise in artificial selection experiments. One explanation for this phenomenon is that traits that are favored under the experimental conditions are negatively correlated with the trait under selection. In this experiment, CH~4~ oxidation rate could be negatively correlated with traits that favor persistence in jars [@conner2003]. This tradeoff would result in a negative heritability because, in the absence of an imposed selection pressure, passaging in jars would favor microbes with lower CH4 oxidation rates.

@@

Even though we observed an increase in methane oxidation rate in the Positive treatment and a difference in composition between the two treatments, we did not observe an increase in the relative abundance of methanotrophs. This contrasts with the usual assumption that the final step in a metabolic pathway is the rate-limiting step (citation?). In certain ecosystems, methane oxidation rate is correlated with the abundance of methanotrophs as estimated from marker genes (Cite. I know that one methanogenesis paper, not sure about methanotrophy). However, it does not appear generally true that the rate of an ecosystem function is limited by the abundance of the gene that encodes the final step in that pathway (cite rocca, maybe some of the methane work.) Since we did not observe an increase in the relative abundance of methanotrophs in the treatment that displayed in increase in methane oxidation rate, we may have to look at other microbial functional groups within an ecosystem to understand what drives changes in the rate of ecosystem functions.

@@

As more is understood about the role of the microbiome in regulating ecosystem functions and the specific taxa involved, more detailed experiments could selectively add or remove important taxa or manipulate their relative abundance to start to estimate the quantitative relationship between specific taxa and the rate of an ecosystem function. Both our differential abundance analysis and this reductionist approach to manipulating individual taxa assumes that taxa act individually on ecosystem function. However, it is much more likely that the microbiome affects the rate of ecosystem function through the interactions of multiple taxa and future analyses with greater replication should test the role of multiple taxa simultaneously using machine learning methods [@topçuoğlu2020]

>>>>

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
any ecosystem function. This challenge remains because it is difficult to
manipulate microbiome composition directly, we cannot know a priori which members of the
microbiome limit the rate of an ecosystem function, and microbiomes covary with
the environment. We propose that artificial ecosystem selection
can overcome these challenges. Artificial ecosystem selection involves creating
replicate ecosystems, selecting replicates with a desired ecosystem-level
trait, and using the microbiomes from the selected ecosystems to inoculate a
new set of ecosystems. We performed artificial ecosystem selection on replicate
soil microcosms to select for increased soil methane oxidation rate. We
observed a strong response to selection with a 51% increase in methane oxidation rate per round over five
rounds of selection. This suggests that a large portion of the variation in
methane oxidation rate in these soils can be attributed to microbiome
variation. We also found that selection did not enrich for known methane
oxidizers; rather microbes that responded to selection included `r length(da_fams)`
families such as *Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result
is in contrast to the typical assumption that the rate of an ecosystem function
is limited by the final step in the associated microbial pathway. Our study
demonstrates that variation in microbiome composition can contribute to
variation in the rate of ecosystem function.

@@

We applied artificial ecosystem selection on soil microbiomes by selecting on
the ecosystem-scale CH~4~ oxidation rate.  Using this approach, we were able to
estimate how much of the variation in CH~4~ oxidation rate was associated with
variation in the microbiome.  To identify which microorganisms could be markers
of CH~4~ oxidation rate, we compared the composition of the artificially
selected microbiomes to a control set of microbiomes without selection.  We
then evaluated whether these markers met our underlying assumptions about
which taxa limit the rate of ecosystem function.

@@

We estimated heritability as the regression of mid-offspring on mid-parent (`r figures("herit", display = "cite")`).
Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
Neutral treatment `r print_lm(herit_model, 2)`.

@@

To identify taxa that responded
to selection on soil CH~4~ oxidation rate, we tested the differential relative
abundance of families in the Positive jars relative to the Neutral jars within
Passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
families that were enriched or depleted in the Positive treatment relative to
the Neutral treatment (`r figures("da", display = "cite")`).

@@

We did not identify any methanotrophic families in the overall differential abundance analysis. 
However, we wanted to look more closely at the known methanotrophs in our dataset to be sure they did not have an effect.
To do this, we subset all of the ASVs in our dataset that were in families that contained methanotrophs.
Only two families were represented: *Methylacidiphilaceae* and *Beijerinckiaceae*. 
Aggregating reads at the family level, neither family was differentially abundant between the two treatments.
However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the Positive treatment and the genus
*Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the
Positive treatment.  While many *Beijerinckiaceae* are methanotrophs, several
taxa in this family have lost the the ability to oxidize CH~4~ and it appears
that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow
on methanol [@tamas2014, @dedysh2016]. Based on this analysis,
it appears that no methanotrophs were enriched in the Positive treatment.

@@

We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome can generate variation in the rate of ecosystem function independent of the underlying environmental conditions. While it is well understood that the microbiome mediates a variety of ecosystem functions, a fundamental question in microbial ecology is whether we need to understand variation in the composition of microbiomes to understand variation in the rate of ecosystem functions. To address that question, we performed artificial ecosystem selection on soil microbiomes by selecting for ecosystems with high methane oxidation rates. We found that there was a significant increase in methane oxidation rate as a response to selection. Each generation of soil ecosystems was created from the same pool of sterile soil substrate and a small inoculum from the previous generation. Therefore, it is unlikely that this response to selection was due to changes in the abiotic conditions of the ecosystem. In addition, the response was only observed in the Positive treatment and not in the Neutral treatment, which shows that this was not a side effect of passaging in jars. Therefore, we conclude that variation in the microbiome can generate variation in ecosystem functions independent of the environment.

@@

@@

Even though we observed an increase in methane oxidation rate in the Positive treatment and a difference in composition between the two treatments, we did not observe an increase in the relative abundance of methanotrophs. This contrasts with the usual assumption that the final step in a metabolic pathway is the rate-limiting step (citation?). In certain ecosystems, methane oxidation rate is correlated with the abundance of methanotrophs as estimated from marker genes (Cite. I know that one methanogenesis paper, not sure about methanotrophy). However, it does not appear generally true that the rate of an ecosystem function is limited by the abundance of the gene that encodes the final step in that pathway (cite rocca, maybe some of the methane work.) Since we did not observe an increase in the relative abundance of methanotrophs in the treatment that displayed an increase in methane oxidation rate, we may have to consider other microbial functional groups within an ecosystem to understand what drives changes in the rate of ecosystem functions.

@@

As more is understood about the role of the microbiome in regulating ecosystem functions and the specific taxa involved, more detailed experiments could selectively add or remove important taxa or manipulate their relative abundance to start to estimate the quantitative relationship between specific taxa and the rate of an ecosystem function. Both our differential abundance analysis and this reductionist approach to manipulating individual taxa assumes that taxa act individually on ecosystem function. However, it is much more likely that the microbiome affects the rate of ecosystem function through the interactions of multiple taxa and future analyses with greater replication should test the role of multiple taxa simultaneously by, for example, using machine learning methods [@topçuoğlu2020]
